---
title: "WUP_04"
author: "Valerie Rodriguez"
date: "Spring 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <-read.csv(file="GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(mydata,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot()
```

In this box plot its seen that both Lt High School and high school participants are likely to end up below the results of participants in the graduate category. This helped us learn that when the boxes are apart from each other if they were to be in the same column this is usually and indicator that the data will be different.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```

The Q1 of graduate school is 10 while the Q3 of LT high school are both lower than the Q1 of graduate. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```

The probability of there being no difference between these different groups is 0.000236. So this means that there is a difference because that probability number is so small. A multiple comparison test would need to be run in order to find out the difference in categories. 

```{r}
a <- TukeyHSD(model1,conf.level=.99)
par(mar = c(3, 15.7, 2, 2))
plot (a,las=2)
```
```
When we ran the tukeyHSD test the graph displayed both LT high school and gradute and highschooll and graduate have statical differences at 95% level.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
